---
title: "Data Cleaning and Visualization"
author: "Simona Rahi"
date: "1/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data and Library Importing

```{r echo=TRUE, message=FALSE, warning=FALSE}
# importing necessary libraries
library(ggplot2)
library(dplyr)
library(naniar)
library(tidyr)
library(reshape2)
library(data.table)
```

```{r}
# reading in the data and making sure to read empty cells as NA 
tr_data <- read.csv('ml_case_training_data.csv', na.strings = c("", "NA"))
tr_output <- read.csv('ml_case_training_output.csv', na.strings = c("", "NA"))
tr_hist <- read.csv('ml_case_training_hist_data.csv', na.strings = c("", "NA"))
```

tr_data includes information on the customers, with the 'id' being the unique key for each customer. This dataframe contains 16096 rows, meaning 16096 customers. 

tr_output also includes 16096 rows and each row indicates whether the corresponding customer (referenced with the id) churned or not.

tr_hist includes historical price information (both fixed price and variable price) over the span of a year (2015, from January till December). We have 193002 rows, which does not seem right because if each of the 16096 customers had price information over a year, that would mean we should have 16096*12=193152 rows, so we will look into this below as well.

## Data Cleaning and Anomaly Detection

```{r}
# checking if customer ID is actually unique as it should be (in tr_data and tr_output)
unique_id <- unique(tr_data$id)
length(unique_id)

unique_id1 <- unique(tr_hist$id)
length(unique_id1)
```
Indeed it is in both. So we do have 16096 customers.

**Let's check the anomaly in the price history**

```{r}
# checking how many unique customers we have
unique_id_hist <- unique(tr_hist$id)
length(unique_id_hist)
```

The number is correct and matches the other datasets. So it seems that some customers do not have price data for the 12 months of 2015.

```{r}
row_per_customer <- tr_hist %>% group_by(id) %>% summarise(count=n()) %>% arrange(count) %>% ungroup() %>% as.data.frame() 

# writing function to check how many customers have less than 12 months reported for price history
anomalies <- c()
for (i in 1:nrow(row_per_customer)){
  
  if (row_per_customer$count[i] < 12) {
  anomalies[i] <- row_per_customer$id[i]
  }
}

length(anomalies)
```

We see that 106 customers have less than 12 months reported for price history. 

#### Types of Columns 

```{r}
# starting with tr_data
sapply(tr_data, class)
```

We need to change dates into date type instead of factor.

```{r}
# changing corresponding column types to Date

tr_data$date_activ <- as.Date(tr_data$date_activ, format = '%Y-%m-%d')
tr_data$date_activ <- as.Date(tr_data$date_end, format = '%Y-%m-%d')
tr_data$date_activ <- as.Date(tr_data$date_first_activ, format = '%Y-%m-%d')
tr_data$date_activ <- as.Date(tr_data$date_modif_prod, format = '%Y-%m-%d')
tr_data$date_activ <- as.Date(tr_data$date_modif_prod, format = '%Y-%m-%d')

```


```{r}
# now on tr_hist
sapply(tr_hist, class)
```

Same issue with the date.

```{r}
tr_hist$price_date <- as.Date(tr_hist$price_date, format = '%Y-%m-%d')
```


```{r}
# finally on tr_output
sapply(tr_output, class)
```

Churn is actually a factor, where 1 indicates yes and 0 indicates no.

```{r}
tr_output$churn <- as.factor(tr_output$churn)
```


#### Missing Values Exploration

We will use a package that helps us easily visualize missing values, it is called naniar. 

```{r}
vis_miss(tr_output)
```

No missing values in the dataframe tr_output.

```{r}
vis_miss(tr_data)
```

In tr_data, we have some columns with missing values. The column campaign_disc_ele is all empty and should be dropped.

```{r}
tr_data <- subset(tr_data, select = -c(campaign_disc_ele))
```


```{r}
vis_miss(tr_hist, warn_large_data = FALSE)
```

In tr_hist, most of the dataset is complete, exactly 99.5% of it. The missing values are all prices so we can replace them with the average price in that column. 

```{r}
# replacing missing value with mean of column
for (i in 3:ncol(tr_hist)){
  tr_hist[is.na(tr_hist[,i]), i] <- mean(tr_hist[,i], na.rm = TRUE)
}
```

#### Shape Transformation

Since the dataframe tr_hist is in long format, we can't really join it on the other datasets if needed. So we will transform it to wide format, so now each row will represent a customer (similar to the others), and we will have a column for each month for each price.

```{r}
# transforming to wide format
wide_tr_hist <- dcast(setDT(tr_hist), id~price_date, value.var=c('price_p1_var', 'price_p2_var', 'price_p3_var', 'price_p1_fix', 'price_p2_fix', 'price_p3_fix'))
```

```{r}
df1 <- merge(tr_data, tr_output, by = 'id')
big_df <- merge(df1, wide_tr_hist, by ='id')
```

```{r}
library(ggcorrplot)
ggcorrplot(big_df, type = 'lower', outline.col = 'black', lab=TRUE, ggtheme = ggplot2::theme_gray, colors = c('#6D9EC1', 'white', '#E46726'))
```


## Data Exploration and Visualization

How many churned and non churned do we have 

```{r}
g <- ggplot(tr_output, aes(churn))
g + geom_bar()
```

Most of the customer we have in the dataset did not churn. There is a large discrepancy between the numbers.

**Summary Statistics**

**Distribution of some parameters**
